---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic <- read_csv("data/titanic.csv")

titanic
```

```{r}
titanic <- 
  titanic %>% 
  mutate(crew = class %in% c("engineering crew", "victualling crew", "restaurant staff", "deck crew"))
```

```{r}
titanic %>% 
  filter(crew == TRUE) %>% 
  summarise(prob = sum(survived == "yes")/n())
```
```{r}
titanic %>% 
  filter(crew == FALSE) %>% 
  summarise(prob = sum(survived == "yes")/n())
```


1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
rolldie(1) %>% 
  filter(X1 < 5)

4/6
```

2. How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
rolldie(3)

216
```

3. In an experiment where we roll four dice:

i. how many outcomes are in the event total_equals_5_6_7 (i.e. the total of the four dice equals 5, 6 or 7)?
ii. what is the probability of this event?
iii. what assumptions do we apply in answering part ii?

```{r}
rolldie(4) #1296


```

