---
title: "R Notebook"
output: html_notebook
---
# Question 1. Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
backpack
dim(backpack)
names(backpack)
```

# Question 2. What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
ggplot(backpack) +
  aes(x = backpack_weight,) +
  geom_bar(fill = "pink")
```


# Question 3. Are there differences in backpack weight between sex?
```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex) +
  geom_bar() +
  facet_wrap(~sex)
```


# Question 4. What is the relationship between body weight and backpack weight?
```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight) +
  geom_point()
```

# Question 5. Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.
```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight, colour = sex) +
  geom_point()
```
The increase in backpack weight doesn't really correlate with sex as the body weight increases. 

Question 6. What proportion of students report back pain?
```{r}
backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>% 
ggplot() +
  aes(x = back_problems) +
  geom_bar() +
scale_x_discrete(labels = c("No", "Yes"))
```


Question 7. Is this proportion different between men and women?
```{r}
backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>% 
ggplot() +
  aes(x = back_problems) +
  geom_bar() +
scale_x_discrete(labels = c("No", "Yes")) +
   facet_wrap(~sex)
```
```


Question 8. Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.
