---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library()
```
# 1.Load the code_clan_tweets.csv data. 

```{r}
tweets <- read_csv("data/code_clan_tweets.csv")
tweets
```

# Find the number of rows, columns, and list all the variable names.

```{r}
dim(tweets)

names(tweets)
```

# 2. Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
tweets %>% 
  filter(favorite_count >= 1, is_quote == FALSE) %>% 
  count() #or summarise(n = n())

#counts total favorites
tweets %>% 
  filter(is_quote == FALSE) %>% 
  summarise(total = sum(favorite_count))
```

# 3. Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}
tweets %>% 
  group_by(source) %>% 
  filter(is_quote == FALSE) %>% 
  summarise(mean_retweets = mean(retweet_count))
```

# 4. Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}
tweets %>% 
  group_by(media_type) %>% 
  count(favorite_count) %>% 
  arrange(desc(favorite_count))
```

# 5. Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).

```{r}
tweets %>% 
  summarise(mean_text_width = mean(display_text_width))
```

# 6. The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
info <- read.csv("data/code_clan_info.csv")
info
join <- left_join(tweets, info, by = "tweet_id")
join
```

#7. From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

```{r}
codeclan_hashtags <- join %>% 
  select(tweet_id, hashtags) %>% 
  mutate(lowercase_hashtag = str_to_lower(hashtags)) %>% 
  select(-hashtags) %>% 
  drop_na(lowercase_hashtag)

#filter(str_to_lower(hashtags) == hashtags)

codeclan_hashtags
```

